Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-09-23 paper/paper.Rmd
rephrased/rewrote spcv_buffer text

Spatial leave-one-out methods start from a single observation in space and create radial (buffer) zones around which define the partitions.

@@

The "buffering" method from the \pkg{blockCV} package is conceptually related to spatial leave-one-out cross-validation.
It comes with optional support for presence/absence data handling, as often used in the field of species distribution modeling [@hijmans2013].
However, by default, the method considers all points for sampling the training and test sets.

In an ecological/species distribution modeling context, the default assumption of a binary response variable is to have *presence/absence* data.
This means all presence and absence values in the response variable are confirmed in their respective status.
The \pkg{blockCV} package implementation also comes with support for *presence/background* data for binary responses.

All other response observations are assumed to be absent without a proof for this assumption.
When setting argument `spDataType = "PB"`, the method only considers the *presence* observations in a CV scenario.
Background data is used for training the model by default unless argument `addBG = FALSE` is set for which only *presence* values not falling within the buffer zone of the respective fold are used for model training.

Due to its similarity to a LOO CV, this method will result in as many train/test evaluations as observations are present.
In the visualization of the first fold in Figure&nbsp;\@ref(fig:buffer-eval) a buffer zone of 1000 m was used without any custom presence/absence treatment of the data.

@@

>>>>

Spatial leave-one-out methods use individual observations in space as test partitions and apply circular buffer or guard zones around around these test points to enforce a minimum prediction distance. 
Leave-one-disc-out resampling modifies this approach to leave out circular regions centered at observation points. 
Note that this straightforward extension of the LOO method does technically not establish a partitioning as used for CV resampling in the traditional sense.

@@

Leave-one-out CV with buffer and several adaptations for species distribution modeling [@hijmans2013] are implemented in the \pkg{blockCV} package as the so-called "buffering" method and integrated into \pkg{mlr3spatiotempcv} under the label `"spcv_buffer"`.
In species distribution modeling, the response variable can either be recorded as presence/absence data or as presence/background information; both options are supported by this implementation. 
By default, the dataset contains confirmed presence and confirmed absence observations, i.e. locations where a species was observed and not observed, respectively, and therefore spatial LOO CV in its usual sense can be carried out.
Figure&nbsp;\@ref(fig:buffer-eval) shows the first test fold generated with this method for presence/absence data with a buffer distance of 1000 m.

@@

In the presence/background (or presence-only) situation, in contrast, only presence observations are recorded, and all other locations within the study area are referred to as the background and considered as pseudo-absences.
Presence/background modeling can be enabled with the argument `spDataType = "PB"`.
In this situation, the method constructs test folds that are centered at the recorded presence locations, offering two different modes of operation.
With `addBG = TRUE` (the default), all background points with a distance of `theRange` around a test (presence) point are included in the test fold as absence data.
The `addBG = FALSE` setting, in contrast, no background data is added to the test fold, which then contains only one (presence) observation, and only the data at a distance of `theRange` or greater are included in the training sample, including background data from these areas.

The application of LOO methods can be computationally expensive since the method cycles through the entire dataset and fits one model for each test fold.
